---
title: "VolBond"
author: "Pieter Pijls"
date: "6 April 2018"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document: default
  word_document: default
graphics: yes
---

# Load packages

```{r, include=FALSE}
rm(list=ls())

packages <- c("data.table","mgcv","evtree", "classInt", "magrittr",
              "rgdal", "ggplot2", "ggmap","grid","gridExtra","xtable","sas7bdat",
              "plyr","knitr","kableExtra","xtable","readxl","xlsx","PerformanceAnalytics",
              "reshape","dplyr","gtable","ggforce","lattice","shinydashboard","rsconnect","rdrop2",
              "plotly", "shiny", "openxlsx","rsconnect")

suppressMessages(packages <- lapply(packages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x)
    library(x, character.only = TRUE)
  }
}))
```

# Import Data

First, I import the swap rate curves from 2002 to 2018.

```{r}
# Import Data (swap rates)
setwd("C:/PIETER/INTERNSHIP P&V/Volbond")
rates <- read.xlsx("./EURIBOR vanilla interest rate swap contract.xlsx",colNames=TRUE,sheet=1)
# create date object
rates$DATE   <- seq(as.Date("2001/1/1"), by = "month", length.out = 206)

rates20 <- as.data.frame(cbind(seq(as.Date("2001/1/1"), by = "month", length.out = 206), rates$`20`))
rates20$V1 <- as.Date(rates20$V1)
```

We plot the evolution of the swap rate of 20Y. 

```{r}
# plot evolution swap rate 20y
g <- ggplot(rates20, aes(x=rates20$V1,y=rates20$V2)) + geom_line() + ggtitle("Evolution swap rate 20Y") +
           xlab("Date") + ylab("Rate")

ggplotly(g)
```

# Volbond data

We import data for the Volbond. The first column is the date, the second is the CMS rate 20Y in advance. The third column corresponds with the CMS 20Y rate in Arrears. The final column is the coupon which is equal to 3.5*absolute_value(CMS20Y in Advance - CMS20Y in Arrears).

```{r}
# Import data for the Volbond
volbond20 <- rates <- read.xlsx("./Volbond20Y.xlsx",colNames=TRUE,sheet=1) 
volbond20$DATE <- seq(as.Date("2001/1/1"), by = "month", length.out = 206)
volbond20 <- volbond20[-1,]
tail(volbond20)
```

We plot the evolution of the CMS rate 20y and the Volbond coupon. 

```{r}
g <- ggplot(volbond20, aes(x=volbond20$DATE)) + geom_line(aes(y=volbond20$Advanced, colour = "CMS20Y")) + geom_line(aes(y=volbond20$Coupon, colour = "Coupon (%)")) + ggtitle("Evolution swap rate 20Y") +
           xlab("Date") + ylab("Rate") + theme_bw()
ggplotly(g)
```

Next we create a histogram of the coupons

```{r}
hist(volbond20$Coupon, breaks=20, xlab = "Coupon (%)")
```

We also have some summary stats. 

```{r}
# summary statistics coupons
summary(volbond20$Coupon)
```

The standard deviation is equal to 

```{r}
# standard deviation
sd(volbond20$Coupon[12:nrow(volbond20)])
```

# Example **without** shout option

In this example we calculate the payoff of the Volbond. We assume that we started investing in the Volbond in January 2001. The first coupon is paid out in January 2002. The last coupon is received in January 2016. We let the shout option outside the analysis.

```{r}
#1# Example without shout option

n = 10000000 # Investment of 10mio
t = 15 # duration bond

coupon <- matrix(ncol = 3,nrow=t)

for (i in 1:t) # nyears(volbond20$DATE)
{
  coupon[i,1] <- as.Date(volbond20$DATE[12*i])
  coupon[i,2] <- volbond20$Coupon[12*i]
  coupon[i,3] <- coupon[i,2]*n
}

as.Date(coupon[,1])
```

The total return of this investment is equal to 24%.

```{r}
# total return
totalreturn <- (sum(coupon[,3])+n)/n
totalreturn
```

The yearly return is equal to 1.479%. 

```{r}
# yearly return
yearlyreturn <- totalreturn^(1/t)-1
yearlyreturn
```

# Example **with** shout option (e.g. lock in rate from 3%)

In this example we incorporate the feature of the shout option. We assume management will lock in the rate when the previous coupon rate was higher as 3%. Again we assume that we started investing in the Volbond in January 2001. The first coupon is paid out in January 2002. The last coupon is received in January 2016.

```{r}
n = 10000000 # Investment of 10mio
t = 15 # duration bond
shout = 0.03 # treshhold when rate is locked in

couponshout <- matrix(ncol = 3,nrow=t)

for (i in 1:t) # nyears(volbond20$DATE)
{
  couponshout[i,1] <- as.Date(volbond20$DATE[12*i], format="%y-%m-%d")
  couponshout[i,2] <- volbond20$Coupon[12*i]
  couponshout[i,3] <- couponshout[i,2]*n
}

# loop to lock in values higher as 3%
for (i in 2:t)
{
if (couponshout[i-1,2]>shout)
  {
    couponshout[i,2] = couponshout[i-1,2]
    couponshout[i,3] = couponshout[i-1,3]
  }
  else
  {
   couponshout[i,2] = couponshout[i,2]
   couponshout[i,3] = couponshout[i,3]
  }
}

colnames(couponshout) <- c("Date","Coupon (%)", "Coupon in ???")
couponshout
```

The total return is higher as we lock in the rate in from January 2005.The coupon is equal to $327.241$ euro when we would invest $10$ mio euro. Therefore the total return is higher compared with the example where we do not use the shout option. When we use the shout option on the right moment we can increase significantly the return on our investment. 

```{r}
# total return
totalreturn <- (sum(couponshout[,3])+n)/n
totalreturn
```

The yearly return is equal to

```{r}
# yearly return
yearlyreturn <- totalreturn^(1/t)-1
yearlyreturn
```

# Different starting dates for the VolBond (without shout option)

In this analysis we use thirteen different starting dates from JAN01,FB01,MAR01,... to FEB02. Therefore, the first column corresponds with the coupon in % when we start investing in January 2001. The second column corresponds with the coupon % when we started investing in February 2001. We calculate the total and yearly return for different staring dates. In this example we do not use the shout option. 

```{r}
n = 10000000 # Investment of 10mio
t = 15 # duration bond

coupon <- matrix(ncol = 13, nrow=t)

# loop over different starting dates
for (j in 1:13)
{

# loop over coupon pay outs
for (i in 1:t) # nyears(volbond20$DATE)
{
  coupon[i,j] <- volbond20$Coupon[(12*i)+j]
}

}  
  
colnames(coupon)<- as.Date(volbond20$DATE[13:25]) 
```

The total returns for the different starting dates from JAN01,FB01,MAR01,... to FEB02 are equal to. The lowest return is 22.9%, the maximum is 33.2%.

```{r}
# total return
totalreturn <- colSums(coupon)
totalreturn
```

```{r}
# yearly return
yearlyreturn <- (1+totalreturn)^(1/t)
yearlyreturn
```

The average yearly return is 1.6%. The maximum return is 2% while the minimum return is 1.4%. 

```{r}
summary(yearlyreturn)
```


